resistanceModelNotesAndy

### 2016-04-16 paper1_results_figs.Rmd
results for a new insecticide added to an old or on its own

```{r,echo=FALSE,results='hide'}
### NEW part to analyse what happens when a new insecticide is added to an existing one
### either to replace it or to continue in a mixture

@@

  indicesI2_lessthan_I1 <- which( resistPointsI1_T$start_freq_allele2 < resistPointsI1_T$start_freq_allele1 )
  #this shows about half, I also checked no equals which is good
  #length(indicesI2_lessthan_I2)
  #[1] 4983

  #NewOnly first make a copy from existing
  resistPointsNewOnly <- resistPointsI1_T
  #replace those values where I2 is the 'new' insecticide with lower resistance
  resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
     resistPointsI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]

  #OldPlusNew first make a copy from existing
  resistPointsOldPlusNew <- resistPointsMixI1_T
  #replace those values where I2 is the 'new' insecticide with lower resistance
  resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
  resistPointsMixI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]

@@

@@

@@

    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
             geom_point(shape=1, colour='red') +
             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
             geom_hline(yintercept = 0, linetype=3) +
             ylim(-1,1) +
             ggtitle(paste("PRCC difference between sequential and mixture adaptive")) +
             xlab(NULL)
          )

>>>>

```{r,echo=FALSE,results='hide'}
### 4/2016 NEW part to analyse what happens when a new insecticide is added to an existing one
### either to replace it or to continue in a mixture

@@

  indicesI2_lessthan_I1 <- which( resistPointsI1_T$start_freq_allele2 < resistPointsI1_T$start_freq_allele1 )
  #this shows about half, I also checked no equals which is good
  #length(indicesI2_lessthan_I2)
  #[1] 4983
  #check whether this works
  #check <- resistPointsI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]
  #aha! problem I was having is that I1 & I2 inputs are not swapped around
  #if I wanted to swap the inputs for I1 & I2 around, would need to do other coefficients etc.
  #much potential confusion
  #might be better just to restrict to the runs in which I2 < I1
  resistPointsNewOnly <- resistPointsI2_T[indicesI2_lessthan_I1, ]
  resistPointsOldPlusNew <- resistPointsMixI2_T[indicesI2_lessthan_I1, ]

  # #this bit now replaced by the 2 previous lines
  # #NewOnly first make a copy from existing
  # resistPointsNewOnly <- resistPointsI1_T
  # #replace those values where I2 is the 'new' insecticide with lower resistance
  # resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
  #    resistPointsI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]  
  # 
  # #OldPlusNew first make a copy from existing
  # resistPointsOldPlusNew <- resistPointsMixI1_T
  # #replace those values where I2 is the 'new' insecticide with lower resistance
  # resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
  # resistPointsMixI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]

@@

  #*****************
  # to check
  # are there any cases where startfreqI1 > startfreqI2
  # by my method above i don't think there should be
  # i think min should be 1
  min(dif_oldnew_newonly$resist_start_1_div_2)
  max(dif_oldnew_newonly$resist_start_1_div_2)

  min( dif_oldnew_newonly$start_freq_allele1 - dif_oldnew_newonly$start_freq_allele2)
  max( dif_oldnew_newonly$start_freq_allele1 - dif_oldnew_newonly$start_freq_allele2)

  #resistPointsNewOnly
  min(resistPointsNewOnly$resist_start_1_div_2)
  max(resistPointsNewOnly$resist_start_1_div_2)

  # because I now just use runs where startfreq12 < I1
  # resist_start_1_div_2 should be the inverse of resist_start_hi_div_lo
  # or at least directly correlated
  # yes they are all on a single curved line
  # plot( resistPointsNewOnly$resist_start_1_div_2 ~ resistPointsNewOnly$resist_start_lo_div_hi)

  #****************
  #not sure I need to do this because all results already show mixture better
  #now I want to subset just those runs where start freq old is > 10 * new

  #proportion of runs in which mix better than sole
  num_mix_better <- length(which(dif_oldnew_newonly["mix_minus_sole0.5"] > 0))
  num_all <- nrow(dif_oldnew_newonly)
  num_mix_better/num_all
  #1
  #?now mix always seems to be better
  #which kind of makes sense, any insecticide better than none

@@

## Fig. xx3 Using a new insecticide on its own or in combination with old

  ggSubset <- dif_oldnew_newonly

    y <- "mix_minus_sole0.5"

    print( ggplot(ggSubset, aes_string(x=i, y=y)) + #, colour="strategy")) + 
             geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
             geom_smooth(linetype='dashed',size=1, colour="red") +
             #facet_wrap( ~ strategy) +
             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
             labs(title = i) +
             geom_hline(yintercept = 0, colour="white")
           )
  }

# Fig xxx PRCC difference in time-to-resistance between sole new and mix old+new
```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}

    x <- dif_oldnew_newonly[,1:num_inputs]
    y <- dif_oldnew_newonly["mix_minus_sole0.5"]

    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
             geom_point(shape=1, colour='red') +
             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
             geom_hline(yintercept = 0, linetype=3) +
             ylim(-1,1) +
             ggtitle(paste("PRCC benefit of mix old+new over sole new (old=1, new=2)")) +
             xlab(NULL)
          )
```

@@

    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
             geom_point(shape=1, colour='red') +
             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
             geom_hline(yintercept = 0, linetype=3) +
             ylim(-1,1) +
             ggtitle(paste("PRCC difference between sequential and mixture adaptive")) +
             xlab(NULL)
          )
```

### 2016-04-16 resistanceModelNotesAndy.Rmd
results for a new insecticide added to an old or on its own

@@

~ check on dif_oldnew_newonly["mix_minus_sole0.5"]

>>>>

BUT this is still for all runs (i.e. not just the ones where start freq of the old is > 10 * the new)

** 11/4/2016 **

I think I want to plot this against each input variable to check whats going on, i.e. where is the mix better than sole use.

dif_oldnew_newonly["mix_minus_sole0.5"]

see Fig xx3

In most (70%) runs mixture are better, but that does leave 30% where sole better.

Eyeballing the plots suggests that exposure is still the most influential input. At high exposures the benefit of the mixture is less.

May be something weird going on with start_freq_allele2, noticeable lack of runs where mix-sole < 0 as start_freq_allele2 gets higher ...

Are the inputs relating to allele1 & 2 still meaningfull with the post-processing that I do ?

No.
aha! problem I was having is that I1 & I2 inputs are not swapped around.

  #might be better just to restrict to the runs in which I2 < I1
  resistPointsNewOnly <- resistPointsI2_T[indicesI2_lessthan_I1, ]
  resistPointsOldPlusNew <- resistPointsMixI2_T[indicesI2_lessthan_I1, ]

when I decided to restrict to runs where startfreqI2 < I1 mix seems to be better all the time ??

? am I doing something wrong ?
seems not

Eyeballing fig xx3

when I1 is the old :
Most important inputs & their effect on the benefit of mixture are :
exposure : reduces benefit of mixture
effectiveness i1,old : increases benefit of mixture
dominance i2,new : reduces benefit of the mixture
selection coefficient i2 : reducces benefit of the mixture

My thoughts on mechanisms.
The mixture does better because the old insecticide kills some mosquitoes that have developed resistance to the new, thus reducing rate of increase in resistance to the new. 
Exposure. Low exposure reduces the benefits of mixtures. This could be because the beneficial effect of the old insecticide is reduced.
Effectiveness of the old insecticide. High effectiveness increases the benefit of the mixture because more of the mosquitoes that have developed resistance to the new insecticide are killed.
Dominance. High dominance of the new insecticide allele decreases the benefit of the mixture. High dominance in resistance to the new insecticide should allow more mosquitoes to escape the new insecticide therefore I would expect the benefit of the mixture to be greater. But its the other way around.

I probably need to get straighter in my head how dominance works.
Does dominance effect the relative performance of the 2 insecticides where mosquitos are exposed to both. i.e. Does high dominance of the new allele mean that a mosquito is more likely to survive the presence of the old insecticide ?

To understand the mechanisms more I could look at the effects of the inputs on each strategy on its own.

These results are consistent with previous ones.

~ ran PRCC on the new results (confirmed same important params)

copied new bits into keep_old_insecticide.Rmd and sent the pdf to Ian

** 12/04/2016 **

Ian was surprised that mixtures better in all runs.

He suggested that The Curtis eg where sole better might be strange and not fall within our parameter space.

~ check whether I can recreate some results from the new analysis by putting the params into the UI, or better by calling runcurtis_f2()

e.g. first run in dif_oldnew_newonly
mix : 180
sole : 48
difference : 132

#a previous example
runcurtis_f2( P_1 = 0.01 , P_2 = 0.3 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )

#putting params of 1st run in
runcurtis_f2( P_1 = 0.0159762192 , P_2 = 0.0076494164 , h.RS1_A0 = 0.660797792 , h.RS2_0B = 0.629114044 , exposure = 0.5582827 , phi.SS1_A0 = 0.9441143 , phi.SS2_0B = 0.9695714 , s.RR1_A0 = 0.1216252 , s.RR2_0B = 0.1720911, addCombinedStrategy=FALSE )

gives warning :
critical point not reached for i1 or i2: cutoff_i1_mix=Inf cutoff_i2_mix=Inf

is this because ?? :
1. I've pasted wrong
no
2. an indexing issue means that inputs and results in the table are misaligned

3. what i'm running in these new runs is different from whats done in the plot
don't think so
4. extra defaults no passed to the plot are different from those used in sensi runs
***aha! could be maleExposureProp and correctMixDeployProp

Looking at the plot sole use does look like it takes about 48 as in the results, but for I2 in the mix looks like it should take ~600 rather than 180 as suggested in the results.

New corrected version :

runcurtis_f2( P_1 = 0.0159762192 , P_2 = 0.0076494164 , h.RS1_A0 = 0.660797792 , h.RS2_0B = 0.629114044 , exposure = 0.5582827 , phi.SS1_A0 = 0.9441143 , phi.SS2_0B = 0.9695714 , s.RR1_A0 = 0.1216252 , s.RR2_0B = 0.1720911, addCombinedStrategy=FALSE, maleExposureProp = 0.90820779, correctMixDeployProp = 0.6008410 )

YES this does now match. Hurrah!
(also shows that maleExposureProp and/or correctMixDeployProp in isolation can have a big effect that I could explore in the mechanisms paper).

Should I do for another result too, just to confirm ?

Hi Ian,
I would like to see if I can recreate Curtis Table 1, (vi).

It makes my head hurt trying to work out what our inputs should be based on his table.

Could you help me out by filling in the below * ? (or is it not possible and I'm missing something?).
# start frequencies of each resistance allele
P_1 = *, 
P_2 = * , 
# dominances
h.RS1_A0 =  *, 
h.RS2_0B =  *, 
# exposures
exposure =  *,
# effectivenesses
phi.SS1_A0 =  *, 
phi.SS2_0B =  *, 
# selection coefficients
s.RR1_A0 =  *, 
s.RR2_0B =  *

Here are the equivalent values used to recreate his figure 2.
# start frequencies of each resistance allele
P_1 = 0.01 , 
P_2 = 0.01 , 
# dominances
h.RS1_A0 = 0.17 , 
h.RS2_0B = 0.0016 , 
# exposures
exposure = 0.9 ,
# effectivenesses
phi.SS1_A0 = 0.73 , 
phi.SS2_0B = 1 , 
# selection coefficients
s.RR1_A0 = 0.23 , 
s.RR2_0B = 0.43

Back from Ian :

P_1 = 0.01,
P_2 = 0.001,
# dominances
h.RS1_A0 =  1,
h.RS2_0B =  1,
# exposures
exposure =  0.9,
# effectivenesses
phi.SS1_A0 =  1,
phi.SS2_0B =  1,
# selection coefficients
s.RR1_A0 =  1,
s.RR2_0B =  1 )

runcurtis_f2( P_1 = 0.01,
P_2 = 0.001,
# dominances
h.RS1_A0 =  1,
h.RS2_0B =  1,
# exposures
exposure =  0.9,
# effectivenesses
phi.SS1_A0 =  1,
phi.SS2_0B =  1,
# selection coefficients
s.RR1_A0 =  1,
s.RR2_0B =  1,
addCombinedStrategy=FALSE )

This seems to generate a very rapid rise in resistance.

and generrates this error in the graph :
Error in xy.coords(x, y) : 'x' and 'y' lengths differ
addCombinedStrategy=FALSE fixes the error

i think I should switch i1 & 2 frequencies

runcurtis_f2( P_1 = 0.001,
P_2 = 0.01,
# dominances
h.RS1_A0 =  1,
h.RS2_0B =  1,
# exposures
exposure =  0.9,
# effectivenesses
phi.SS1_A0 =  1,
phi.SS2_0B =  1,
# selection coefficients
s.RR1_A0 =  1,
s.RR2_0B =  1,
addCombinedStrategy=FALSE,
addStrategyLabels = FALSE )

This seems to show that resistance to the new insecticide (red) rises faster when alone (dotted) than when in combination (solid) with an old insecticide at a higher frequency. This is contrary to what Curtis said.

Also note that resistance arises much faster than in our sensitivity runs. I think  this is because the selection coefficients (1) are much higher than the range we use (0.1-0.45).

** 15/4/16 **

from Ian :
I have re-read Curtis again and its that paragraph on page 261 that is causing all the problems i.e.

“Example vi shows the, at first unexpected……”

I guess the key phrase  is what he means by “more rapid”…I assume he means more rapid than using it on its own. But that doesn’t tally with your figure. His example is only a 1-genertion argument so I should be able to do some simple arithmetic. *Can you tell me what the allele frequencies were at generation 2 of your plot* and I’ll check against my arithmetic.

to get allele freqs at gen 2 :

tst$results[[2]][1:8,]

@@

~ how to recreate Curtis Table 1, (vi)'the use of a mixture when the initial gene frequencies are unequal leads to a more rapid increase in the frequency of the rarer of the genes'.

~ remember this is still for all runs (i.e. not just the ones where start freq of the old is > 10 * the new)

~ **check this from above**
~ I probably need to get straighter in my head how dominance works.
Does dominance effect the relative performance of the 2 insecticides where mosquitos are exposed to both. i.e. Does high dominance of the new allele mean that a mosquito is more likely to survive the presence of the old insecticide ?

To understand the mechanisms more I could look at the effects of the inputs on each strategy on its own.
